Effectiveness of community-based marine reserves in small-scale fisheries

### 2018-08-16 docs/paper/Villasenor-Derbez.tex
Incorporated Fio's comments

\begin{document}
\onecolumn
\firstpage{1}

\title[Community-based marine reserves: 2018-07-05]{Effectiveness of community-based marine reserves in small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries. Mexico has seen a proliferation in
their implementation of, with 45 TURF-reserves implemented in the last 4
years after a new regulation was published in 2014. However, their
effectiveness has not been formally evaluated accounting for the
intricate social-ecological dimensions and their governance context.
This work combines causal inference techniques and the social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. We identify that
while reserves don't achieve their stated goals of increasing lobster
densities, they continue to receive significant support from the fishing
communities. Our triple-bottom evaluation suggests that the lack of
effect is a combination of factors. First, the lobster fisheries are
already well managed, and it is unlikely that reserves might have a
detectable effect. Second, reserves are not big enough to protect
lobster's home range. Third, some of these reserves might be too young
for the effects to show. However, we argue that these are not failed
enterprises, since reserves can still serve as an insurance mechanism
against errors and uncertainty in management, as well as environmental
variation. Furthermore, these reserves have shaped small-scale fisher's
way of thinking about marine reserves, which can provide solid grounds
to implement more or larger marine reserves.

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
artisanal fisheries face great challenges since they tend to be hard to
monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MR) can greatly improve the performance of coastal
fisheries and the health of the local resouces
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintainance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix}, little empirical evidence
evidence exists of their effectiveness and the drivers of their success
\citep{afflerbach_2014-HP,lester_2017}.

The performance of these systems depends on how environmental and social
factors work combined. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, richness, and densities of organisms within
the protected regions, climate change mitigation, and protection from
environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on
the relationship between socioeconomic and governance structures and
their relationship to reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective.

TURF-reserves can be created as community-based marine reserves. This
bottom-up approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
Community-based spatial closures occur in other places, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
these are difficult to enforce and fishers rely on the exclusive access
granted by the TURF. In an effort to bridge this normative gap, Civil
Society Organizations (CSOs) served as the link between fishers and
government, and set out to create a legal framework that solve this
governance issue. In 2014, a new norm was created, allowing fishers to
request the legal recognition of a community-based reserve under the
name of ``Fish Refuge'' \citep{nom}. These can be implemented as
temporal or partial reserves, which can protect one, some, or all
resources within them. Since then, 45 of community-based marine reserves
along the Pacific, Gulf of California, and Mexican Caribbean coastlines
have gained legal recognition, but their effectiveness has not been
reported in the scientific literature.

This work combines causal inference techniques and the social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. The objective of
this work is twofold. First, provide a triple bottom line evaluation of
the effectiveness of community-based marine reserves that can inform
similar processes in other countries. And second, evaluate the
effectiveness of TURF-reserves established as Fishing Refugia in Mexico
to identify opportunities where improvement or adjustment might lead to
increased effectiveness. On both cases, we draw from the lessons learned
and provide management recommendations to maximize the effectiveness of
community-based marine reserves in small-scale fisheries.

@@

We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
first one was created by the \emph{Buzos y Pescadores de la Baja
California} fishing cooperative, located in Isla Natividad in the Baja
Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
spiny lobster (\emph{Panulirus interruptus}), but other resources like
finfish, sea cucumber, read sea urchin, snail, and abalone are also an
important source of income. In 2006, the community decided to implement
two marine reserves within their fishing grounds to protect commercially
important invertebrate species; mainly lobster and abalone. While these
reserves obtained legal recognition until 2018, they have been well
enforced since their implementation.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). Maria Elena is a
fishing camp --visited intermittently during the fishing season--
belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
Punta Herrero is home to the \emph{SCPP José María Azcorra} cooperative.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves are legally
recognized as Fishing Refugia since their creation.

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species in the reserves and control
sites \citep{fulton_2018}. Size structures are also collected during
fish surveys. We define control sites as regions with habitat
characteristics similar to the corresponding reserves, and that
presumably had a similar probability of being selected as reserves
during the design phase. We focus our evaluation on sites where data are
available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves. Table
\ref{table:com_sum} shows a summary of the TURF-reserves included in
this study.

@@

\begin{table}

@@

\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l|l}
\hline
Category & Indicador & Units\\
\hline
Biological & Lobster density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
Biological & Fish density & org $\mathrm{m}^{-2}$\\
\hline
Socioeconomic & Income from target species & M MXP\\
\hline
Socioeconomic & Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

@@

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfacutals that allow us to control for unobserved
time-invariants. Each ``treated'' community (Isla Natividad and Maria
Elena) has three counterfactual communities.

@@

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
this case the \(Treated\) dummy variable indicates if the community has
a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
captures community-level fixed-effects. These regressions allows us to
make a causal link between the implementation of marine reserves and the
observed trends by accounting for temporal and spatial dynamics
\citep{depalma_2018}. The effect of the reserve is captured by the
\(\lambda_t\) coefficient, and represents the difference observed
between the control site before the implementation of the reserve and
the treated sites at time \(t\) after controlling for other time and
space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
All model coefficients were estimated via ordinary least-squares and
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
Data and code are available on
\href{https://github.com/jcvdav/ReserveEffect}{github.com}.

@@

\subsection{Biological}\label{biological}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). The rapid increase observed for changes in lobster
densities for Isla Natividad on the sixth year (\emph{i.e.} 2012) occur
a year after the hypoxia events described by \citet{micheli_2012-EU}
caused mass mortality of organisms. Likewise, no changes were detected
in fish biomass or invertebrate and fish densities
(\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Full tables with model coefficients are presented
in the supplementary materials (\textbf{S2 Table}, \textbf{S3 Table},
\textbf{S4 Table}).

\subsection{Socioeconomic}\label{socioeconomic}

@@

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for lobster
densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
densities (C), and fish densities (D). Plots are ordered by survey type
(left column: invertebrates; right column: fish). Points are jittered
hotizontally to avoid overplotting. Points indicate the effect size, and
errorbars standard errors. Years have been centered to year of
implementation.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points indicate the
effect size, and errorbars standard errors. Years have been centered to
year of implementation.}
\end{figure}

@@

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators other than the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. The lack of evidence of the effectiveness of
these reserves is surprising since most of the communities show a
positive perception about their performance and continue to support
their presence \citep{ayer_2018}. Analyzing the shortcomings of our
study and understanding the social-ecological context in which these
communities and their reserves operate might provide insights to this
question.

Some works evaluate marine reserves by performing inside-outside
\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
before-after comparisons \citep{betti_2017-lq}. The first approach does
not address temporal variability, and the second can not distinguish
between the temporal trends in a reserve and the entire system
\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
changes provides a more robust measure of reserve effectiveness.
However, this method assumes control sites are a plausible
counterfactual for treated sites. This implies that treated sites would
have followed the same trend as control sites, had the reserves not been
implemented. Nonetheless, overall trends for each site don't show any
significant increases, supporting our findings of lack of change in the
indicators used (\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4
Figure}, \textbf{S5 Figure}, \textbf{S6 Figure}).

A first possible explanation for the lack of effectivenes may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO}. Isla Natividad has the oldest reserve, and our
SES analysis suggests that all communities have a well-established
community-based enforcement system. With these characteristics, one
would expect the reserves to be effective. Maria Elena and Punta Herrero
are relatively young reserves (\emph{i.e.} \textless{} 5 years old);
other community-based marine reserves in tropical ecosystems may take up
to six years to show a spillover effect \citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to reserves being too small. Previous research has shown that
reserves in Isla Natividad yield fishery benefits for the abalone
fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
and perhaps the reserves provide enough protection to these sesile
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2006}. Both lobster fisheries were, at
some point, certified by the Marine Stewardship Council
\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
via species-specific minimum catch sizes, seasonal closures, protection
of ``berried'' females, and escapement windows where traps are allowed
\citep{dof_website_1993}. It is uncertain whether such a well-managed
fishery will experience additional benefits from marine reserves.
Additionally, \citet{gelcich_2008} has shown that TURFs alone can have
greater biomass and richness than areas operating under open acces.
These increased attributes perhaps minimize the difference between TURF
and reserve. Further research should focus on evaluating sites in the
reserve, TURF, and open access areas.

While reserves fail to provide fishery benefits, there are a number of
additional ecological, fisheries, and social benefits. Marine reserves
provide protection to a wider range of species and vulnerable habitat,
like coral reefs. These sites can serve as an insurance against
uncertainty and errors in fisheries management, as well as environmental
shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
serve as a way to compensate for future declines associated to
environmental variation \citep{finkbeiner_2018}. Embarking in a marine
conservation project can bring the community together, which promotes
social cohesion and builds social capital. Furthermore, showing
commitment to marine conservation and sustainable fishing practices
allows fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}.

>>>>

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\onecolumn
\firstpage{1}

\title[Community-based marine reserves: 2018-08-16]{Effectiveness of community-based marine reserves in small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 4
years, 45 TURF reserves have been implemented in Mexico, but their
effectiveness has not been formally evaluated. We combine causal
inference techniques and a social-ecological systems framework to
provide a holistic evaluation of community-based TURF reserves in three
coastal communities in Mexico. We find that while reserves have not yet
achieved their stated goal of increasing lobster densities, they
continue to receive significant support from the fishing communities. A
lack of ecological and socioeconomic effects or reserves likely results
from a combination of factors. First, the lobster fisheries are already
well managed, and it is unlikely that reserves might have a detectable
effect. Second, reserves are not large enough to protect lobster's home
ranges. Third, some of these reserves might be too young for the effects
to show. However, these reserves have shaped small-scale fishers' way of
thinking about marine management, which can provide a foundation for
establishing additional, larger marine reserves needed to effectively
conserve mobile species.

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
artisanal fisheries face great challenges since they tend to be hard to
monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MR) can greatly improve the performance of coastal
fisheries and the health of the local resouces
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintainance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix}, there is little empirical
evidence of their effectiveness and the drivers of their success
\citep{afflerbach_2014-HP,lester_2017}.

The performance of these systems depends on how environmental and social
factors combine and interact. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, species richness, and densities of organisms
within the protected regions, climate change mitigation, and protection
from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1}
\href{http://dx.doi.org/10.1098/rstb.2014.0276}{\textbf{add De Leo and
Micheli 2015}}. Other research has focused on the relationship between
socioeconomic and governance structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective. This is especially important in
social-ecological coastal systems dominated by close interaction and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
no-take regulations are difficult to enforce and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Civil Society Organizations (CSOs) have served as a link
between fishers and government, and set out to create a legal framework
that solve this governance issue. In Mexico, a new norm was created in
2014 allowing fishers to request the legal recognition of
community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
Pesquero} ;\citet{nom}). Fish refuges can be implemented as temporal or
partial reserves, which can protect one, some, or all resources within
their boundaries. Since 2014, 45 of community-based marine reserves
along the Pacific, Gulf of California, and Mexican Caribbean coastlines
have gained legal recognition. However, their effectiveness has not yet
been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and a social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. These three case
studies span a range of ecological and social conditions representative
of different regions of Mexico. The objective of this work is twofold.
First, to provide a triple bottom line evaluation of the effectiveness
of community-based marine reserves that can inform similar processes in
other countries. Second, to evaluate the effectiveness of TURF-reserves
established as Fish Refuges in Mexico to identify opportunities where
improvement or adjustment might lead to increased effectiveness. We draw
from lessons learned in these three case studies and provide management
recommendations to maximize the effectiveness of community-based marine
reserves in small-scale fisheries in Mexico and in other regions around
the world that are using this tool to manage and rebuild their coastal
fisheries.

@@

We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
first one was created by the \emph{Buzos y Pescadores de la Baja
California} fishing cooperative, located in Isla Natividad in the Baja
Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
spiny lobster (\emph{Panulirus interruptus}), but other resources like
finfish, sea cucumber, read sea urchin, snail, and abalone are also an
important source of income. In 2006, the community decided to implement
two marine reserves within their fishing grounds to protect commercially
important invertebrate species; mainly lobster and abalone. While these
reserves obtained legal recognition only in 2018, they have been well
enforced since their implementation.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
cooperative, and similar to Isla Natividad hosts a local community.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves are legally
recognized as Fishing Refugia since their creation.

These communities are representative of their region in terms of
ecology, socioeconomic, and governance aspects. Isla Natividad, for
example, is part of a greater group of fishing cooperatives that
assamble into a Federation of Cooperatives (\textbf{FEDECOOP}). This
group of X communities has been identified as a cohesive group that
often cooperate to better manage their resources
\citep{mccay_2017-1m,mccay_2014-CN} (\textbf{Aceves, 2018?}). All these
communities occur in the sourthenmost raange distribution of Giant Kelp,
and are still under the influence of the California Current System.
Likewise, Maria Elena and Punta Herrero are representative of fishing
cooperatives in the Mexican Caribbean. Together, these three communities
provide an accurate representation of all other fishing communities in
Mexico. While each region has more communities that have established
community-based TURF-reserves, available data would not allow us to
perform the in-depth analysis that we undertake. Yet, given the
similarities among communities and the socioeconomic and governance
setting under which they opperate, it is sage to generalize our results
to other communities in Mexico and other regions around the world.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30x 2 m each) at depths 5-20 m in the reserves and control sites
\citep{fulton_2018}. Size structures are also collected during fish
surveys. We define control sites as regions with habitat characteristics
similar to the corresponding reserves, and that presumably had a similar
probability of being selected as reserves during the design phase. We
focus our evaluation on sites where data are available for reserve and
control sites, before and after the implementation of the reserve. This
provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
sampling design that allows us to capture and control for temporal and
spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
causal inference techniques have proven effective to evaluate marine
reserves, as they allow us to causally attribute observed changes to the
intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
were surveyed annually, and at least once before implementation of the
reserves. Table \ref{table:com_sum} shows a summary of the TURF-reserves
included in this study.

@@

\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l|l}
\hline
Category & Indicator & Units\\
\hline
Biological & Lobster density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
Biological & Fish density & org $\mathrm{m}^{-2}$\\
\hline
Socioeconomic & Income from target species & M MXP\\
\hline
Socioeconomic & Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

@@

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfactuals that allow us to control for unobserved
time-invariants. Each focal community (Isla Natividad and Maria Elena)
has three counterfactual communities.

@@

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
this case the \(Treated\) dummy variable indicates if the community has
a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
captures community-level fixed-effects. These regression models allow us
to establish a causal link between the implementation of marine reserves
and the observed trends by accounting for temporal and spatial dynamics
\citep{depalma_2018}. The effect of the reserve is captured by the
\(\lambda_t\) coefficient, and represents the difference observed
between the control site before the implementation of the reserve and
the treated sites at time \(t\) after controlling for other time and
space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
All model coefficients were estimated via ordinary least-squares and
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
Data and code are available on
\href{https://github.com/jcvdav/ReserveEffect}{github.com}.

@@

\subsection{Biological effects}\label{biological-effects}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
sizes oscillated around zero without clear trends. Full tables with
model coefficients are presented in the supplementary materials
(\textbf{S2 Table}, \textbf{S3 Table}, \textbf{S4 Table}).

\subsection{Socioeconomic effects}\label{socioeconomic-effects}

@@

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for lobster
densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
densities (C), and fish densities (D). Plots are ordered by survey type
(left column: invertebrates; right column: fish). Points are jittered
hotizontally to avoid overplotting. Points indicate the effect size and
standard errors. Years have been centered to year of implementation.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points indicate the
effect size and standard errors. Years have been centered to year of
implementation.}
\end{figure}

@@

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators other than the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. The coastal ecosystems where these reserves are
located have been profoundly affected by climatic and oceanographic
extremes, including warming events, extreme storms and prolonged hypoxia
\citep{micheli_2012-EU}, (\textbf{Beas et al. in prep., Woodson et al in
review}). Despite the lack of evidence of the effectiveness of these
reserves, most of the communities show a positive perception about their
performance and continue to support their presence \citep{ayer_2018}.
Understanding the social-ecological context in which these communities
and their reserves operate might provide insights to this.

Some works evaluate marine reserves by performing inside-outside
\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
before-after comparisons \citep{betti_2017-lq}. The first approach does
not address temporal variability, and the second can not distinguish
between the temporal trends in a reserve and the entire system
\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
changes provides a more robust measure of reserve effectiveness. For
example, we capture previously described patterns like the rapid
increase observed for changes in lobster densities for Isla Natividad on
the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
year after the hypoxia events described by \citet{micheli_2012-EU},
which caused mass mortality of sedentary organisms such as abalone and
sea urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
overall trends for each site don't show any significant increases,
supporting our findings of lack of change in the indicators used
(\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4 Figure}, \textbf{S5
Figure}, \textbf{S6 Figure}).

A first possible explanation for the lack of effectivenes may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO} \textbf{Babcock PNAS 2010}. Isla Natividad has the
oldest reserves, and our SES analysis suggests that all communities have
a well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 5 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; other community-based marine reserves in tropical
ecosystems may take up to six years to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve dsign (\emph{sensu}
\citet{rees_2018}. Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2006}. Both lobster fisheries were
certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
Additionally, lobster fisheries are managed via species-specific minimum
catch sizes, seasonal closures, protection of ``berried'' females, and
escapement windows where traps are allowed \citep{dof_website_1993}. It
is uncertain whether such a well-managed fishery will experience
additional benefits from marine reserves. Additionally,
\citet{gelcich_2008} has shown that TURFs alone can have greater biomass
and richness than areas operating under open acces. These increased
attributes perhaps minimize the difference between TURF and reserve.
Further research should focus on evaluating sites in the reserve, TURF,
and open access areas.

While reserves have failed to provide fishery benefits up to now, there
are a number of additional ecological, fisheries, and social benefits.
Marine reserves provide protection to a wider range of species and
vulnerable habitat, like coral reefs. These sites can serve as an
insurance against uncertainty and errors in fisheries management, as
well as environmental shocks
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{, Aalto et
al, in press Am Nat}). Self-regulation of fishing effort (\emph{i.e.}
reduction in harvest) can serve as a way to compensate for future
declines associated to environmental variation \citep{finkbeiner_2018}.
Embarking in a marine conservation project can bring the community
together, which promotes social cohesion and builds social capital.
Furthermore, showing commitment to marine conservation and sustainable
fishing practices allows fishers to have greater bargaining power and
leverage over fisheries management \citep{prezramrez_2012}.
